---
title: "Rolling Housing Sales for NYC"
author: "Mustafa Sakarwala"
date: "February 8, 2017"
output: html_document
---
```{r setup, include=FALSE}
setwd("C:\\SMU\\Data Science\\Unit5\\R_rollingsales_brooklyn\\data")
#library(plyr)
#library(gdata)
setwd("C:\\SMU\\Data Science\\Unit5\\R_rollingsales_brooklyn\\data")
getwd()
## You need a perl interpreter to do this on Windows.
## It's automatic in Mac
#bk <- read.xls("rollingsales_brooklyn.xls",pattern="BOROUGH")

# So, save the file as a csv and use read.csv instead
bk <- read.csv("rollingsales_brooklyn.csv",skip=4,header=TRUE)


## Check the data
head(bk)
summary(bk)
str(bk) # Very handy function!

## clean/format the data with regular expressions
## More on these later. For now, know that the
## pattern "[^[:digit:]]" refers to members of the variable name that
## start with digits. We use the gsub command to replace them with a blank space.
# We create a new variable that is a "clean' version of sale.price.
# And sale.price.n is numeric, not a factor.
bk$SALE.PRICE.N <- as.numeric(gsub("[^[:digit:]]","", bk$SALE.PRICE))
count <- function(x){
  length(which(x))
}
count(is.na(bk$SALE.PRICE.N))


names(bk) <- tolower(names(bk)) # make all variable names lower case

## TODO: Get rid of leading digits bk$gross.square.feet as above bk$SALE.PRICE
bk$gross.sqft <- as.numeric(gsub("[^[:digit:]]","", bk$gross.square.feet))

# TODO: Get rid of leading digits of as above bk$SALE.PRICE
bk$land.sqft <- as.numeric(gsub("[^[:digit:]]","", bk$land.square.feet))
  
bk$year.built <- as.numeric(as.character(bk$year.built))


## keep only the actual sales
bk.sale <- bk[bk$sale.price.n!=0,]
## for now, let's look at 1-, 2-, and 3-family homes
bk.homes <- bk.sale[which(grepl("FAMILY",bk.sale$building.class.category)),]
```
# TODO: complete plot() with log10 of bk.homes$gross.sqft,bk.homes$sale.price.n
#   as above "bk.sale"

## Analysis of Rolling Housing Sales for NYC
```{r bk0, echo=FALSE}
plot(log(bk.homes$gross.sqft),log(bk.homes$sale.price.n))

## remove outliers that seem like they weren't actual sales
bk.homes$outliers <- (log10(bk.homes$sale.price.n) <=5) + 0

# TODO: find out homes that meets bk.homes$outliers==0
bk.homes <- bk.homes[which(bk.homes$outliers==0),]
```



#By plotting the data for Family houses, it is observed there is an exponential relationship between sale price and gross sqft. As gross sqft increases exponentially, sale price does as well. The plots of log data also shows clear indication that the original data has many outliers. 

#After removing the outliers and, the plotted data to base 10 log below looks more scattered. 
```{r bk1, echo=FALSE}
plot(log10(bk.homes$gross.sqft),log10(bk.homes$sale.price.n))
```

